The Calculus of Linear Constructions

### 2021-10-17 writing/lics22/lics22.tex
writing
lics22

% document layout
\documentclass[sigplan,screen,review,authordraft]{acmart}
\usepackage[utf8]{inputenc}

% packages
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mdwtab}
\usepackage{syntax}
\usepackage{stmaryrd}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{float}
\usepackage{tikz-cd}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

@@

  \subsubsection{Weakening} \label{weakening}
  CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.

@@

@@

\bibliographystyle{acm}
\bibliography{ref}

>>>>

% document layout
%\documentclass[sigplan,screen,review,authordraft]{acmart}
\documentclass[sigplan,screen]{acmart}
\usepackage[utf8]{inputenc}

% packages
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mdwtab}
\usepackage{syntax}
\usepackage{stmaryrd}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{lstlangcoq}
\usepackage{float}
\usepackage{tikz-cd}

\lstset{basicstyle=\scriptsize,language=Coq,showstringspaces=false}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

@@

  \subsubsection{Weakening} \label{weakening}ß
  CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.

@@

  \section{Extensions}
  \subsection{Inductive Types}
  Readers familiar with the literature on linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective. Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written this way will be incredibly unwieldy.

  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic}. The presence of universe sorts $U$ and $L$ allows users to define inductive types to be non-linear or linear. The lack of ! exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching. The $\otimes$ connective becomes a definable inductive type.

  \begin{figure}[h]
  \vspace{-1.4em}
  \caption{Inductive Type Definition and Elimination}
  \begin{lstlisting}
  (* Sigma Definition *)
  Inductive Sigma (A: Type) (F: A -> Type): Type :=
  | pair : (x: A) -> F x -> Sigma A F.

  (* Tensor Definition *)
  Inductive Tensor (A: Linear) (B: Linear): Linear :=
  | tpair: A -> B -> Tensor A B.

  (* FTensor Definition *)
  Inductive FTensor (A: Type) (F: A -> Linear): Linear :=
  | fpair: (x: A) -> F x -> FTensor A F.

  (* Sigma Elimination
     x and y are unrestricted in clause body *)
  match m with
  | pair x y => (* clause body *)
  end

  (* Tensor Elimination
     x and y must be used once in clause body *)
  match m with
  | tpair x y => (* clause body *)
  end

  (* FTensor Elimination
     x is unrestricted in clause body, 
     y must be used once in clause body *)
  match m with
  | fpair x y => (* clause body *)
  end
  \end{lstlisting}
  \label{inductive}
  \Description{}
  \end{figure}

  Figure \ref{inductive} is an excerpt taken from our implementation's prelude, with syntax drawing heavily from Coq.

  \begin{itemize}
    \item \texttt{Sigma} defines the standard dependent sum type. Since first component of the pair constructor is of non-linear type, the type of the second component is allowed to depend on the first component.
    \item \texttt{Tensor} defines the $\otimes$ connective for CLC. Both components of the \texttt{tpair} constructor are of linear types. When a \texttt{Tensor} term is eliminated through pattern-matching, its first and second components must be used exactly once in the body of its pattern clause.
    \item \texttt{FTensor} fulfills a role similar to Krishnaswami et al.'s $F$ adjoint connective \cite{neel15}. The type of \texttt{FTensor}'s first component is non-linear. The linear type of its second component is allowed to depend on the first component. Due to the fact that \texttt{FTensor}'s second component is linear, \texttt{FTensor} must be linear as well, otherwise duplication of restricted variables may occur.
  \end{itemize}

  Due to the complexity of checking soundness for dependent linear type definitions (such as \texttt{FTensor}) and pattern-matching based elimination, we have not verified the algorithm employed by our implementation yet.

@@

\bibliographystyle{acm}
\bibliography{../ref}
